CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-10-01 PaperRMFI2019/samplepaper.tex
small fixes

\begin{document}
%
\title{CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate chest, waist and 
pelvis circumference from 3D human body meshes. Our motivation is to use 
this data as ground truth for learning with convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
$\textit{manually}$ these anthropometrical measurements directly from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast our method can be used on 
3D meshes automatically. We conduct two experiments. 
In the first experiment we synthesize 10 human body meshes. Then we apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively. We observe that the measurements can indeed be used 
to estimate the shape of a person. The second experiment asses the 
plausibility of our approach where we use the calculated human dimensions as 
a ground truth to train a small CNN. After 
having trained the network with 
our data, we demonstrate that learning converges, achieving $x$ percent 
prediction error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{subjects_2_6_calvis_segmented.png}
	\end{center}
	\caption{Mesh segmentation in chest (red), waist (green) and pelvic (blue) 
	regions. We show male (left) and female (right) subjects in frontal and 
	lateral view.}
	\label{fig:axilla-recognition}
\end{figure}

@@

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point in the mesh at the proper location under 
the arms. One point suffices because we require only a reference along the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the left shoulder joint $j_{ls} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

@@

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex with minimum $x$,$y$ and $z$ 
coordinates, respectively. Distance $d(v^z_{max},v^z_{min})$ is the euclidean 
distance between vertices with maximum and minimum $z$ coordinate (figure x).

@@

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{axilla_recognition_final.png}
	\end{center}
	\caption{Axilla recognition based on the skeleton joint. We cast a ray from 
	the left shoulder joint in direction to the middle left 
		inferior edge of the bounding box (left). The ray intersects the mesh 
		at point $p_a$ that we consider the axilla center. We plot the 10 
		nearest vertices to $p_a$ (right). As it can be observed, the axilla is 
		properly recognized.}
	\label{fig:axilla-recognition}
\end{figure}

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference bellow the arms 
at the \textit{widest} part of the torso and the waist circumference at the 
\textit{narrowest} part after the chest but above the hips. This intuition is 
compliant 
with prior body dimensions standardized definition, for example, the European 
standard EN 13402-1 \cite{en13402-1}. Similarly, the pelvis 
circumference is measured often around the widest part of hips and 
buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at a point $q \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the intersection boundary length as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters until we reach the bounding box bottom 
$q_b$. Because the mesh $\mathcal{M}$ and the plane normal $\vec{n}_{||}$ 
remains the same we can drop them from the notation. For every slice at point 
$q$, we compute $\mathcal{BL}(q)$.

\begin{align}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ \mathcal{BL}(q_t), 
\mathcal{BL}(q_t+m), \cdots, \mathcal{BL}(q_b) \right]
\end{align}

Finally, we can define the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ that maps every slice index $k \in \{0, 1, \cdots, 
|\mathcal{L}|-1\}$ to the corresponding boundary length $\mathcal{BL}(q)$ in 
the mesh slide vector $\vec{\mathcal{L}}$.

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(k) = \mathcal{BL}(q)
\end{align}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. Note that the function as 
defined by equation \ref{eq:mesh_signature} is bounded and not continuous. It 
resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

In general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. However, there is no consensus in the 
literature regarding the method to calculate chest, waist and hip 
circumference. While Yansel et al., 2020 define waist circumference as the 
blalal. That is why we first establish the regions and the search for local 
extrema in these regions.

More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$. A horizontal line at height proportion $h$ of the peak 
	intersects 
	it at two points left $p_l$ and right $p_r$.
	\item The plane $\boldsymbol{\pi}$ intersects the mesh $\mathcal{M}$ at the 
	point where the left hip is located defining a path. The length of this 
	path is $P_{lh}$.	
	\item The local maximum $M_{pc}$ with largest value other than $M_g$ 
	immediately prior to $P_{lh}$ represents the pelvis circumference.	
	\item The local minimum $M_{wc}$ posterior to $p_r$ but prior to $M_{pc}$ 
	represents waist circumference.
	\item The local maximum $M_{cc}$ posterior to $p_r$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value in the 
	interval $(p_l; M_{wc})$.
\end{enumerate}

\subsection{Recognition by interval search}\label{subsec:interval_search}

>>>>

\begin{document}
%
\title{CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate chest, waist and 
pelvis circumference from 3D human body meshes. Our motivation is to use 
this data as ground truth for learning with convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
$\textit{manually}$ these anthropometrical measurements directly from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast our method can be used on 
3D meshes automatically. We conduct two experiments. 
In the first experiment we synthesize 10 human body meshes. Then we apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively. We observe that the measurements can indeed be used 
to estimate the shape of a person. The second experiment asses the 
plausibility of our approach where we use the calculated human dimensions as 
a ground truth to train a small CNN. After 
having trained the network with 
our data, we demonstrate that learning converges, achieving $x$ percent 
prediction error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

\begin{figure}[t]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{subjects_2_6_calvis_segmented.png}
	\end{center}
	\caption{Mesh segmentation in chest (red), waist (green) and pelvic (blue) 
	regions. We show male (left) and female (right) subjects in frontal and 
	lateral view. Note how the regions are correctly segmented for different 
	gender and body shapes.}
	\label{fig:mesh_segmented}
\end{figure}

@@

Figure \ref{fig:mesh_segmented} shows the result of applying these equations on 
two meshes. Note how the regions are correctly segmented for different gender 
and body shapes. For example, the male subject on the left is smaller than the 
female subject on the right. Additionally, the subjects have very different 
muscular and fat build.

\subsubsection{Axilla recognition}\label{subsec:armpit_recog}
As mentioned above, we would like to measure the chest circumference below the 
armpit, also known as axilla. This raises the question whether we can recognize 
the axilla. What we want is a point $p_a$ in the mesh at the proper location 
under 
the arms. One point suffices because we require only a reference $p^y_a$ along 
the 
y-axis to define the chest region according to equation \ref{eq:cr}. Therefore, 
we focus on the left shoulder joint $j_{ls} \in \mathbf{J}$. Since the axilla 
is related to 
the shoulder joint, we can cast a ray $\boldsymbol{\iota}$ from it in direction 
to the middle left 
inferior edge of the bounding box outside the mesh at point

@@

Here $v^x_{min}, v^y_{min}, v^z_{min}$ are vertex minimum $x$,$y$ and $z$ 
coordinates, respectively; analog is $v^z_{max}$ the maximum vertex $z$ 
coordinate. The ray 
$\iota$ is then determined by points $j_{ls}$ (inside the mesh) and $p_b$ 
(outside the mesh). Therefore, the ray $\iota$ intersects the mesh at one point 
$p_a$ (Figure \ref{fig:axilla-recognition}), which we consider the axilla 
center.

@@

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{axilla_recognition_final.png}
	\end{center}
	\caption{Axilla recognition based on the skeleton joint. We cast a ray from 
	the left shoulder joint $j_{ls}$ in direction to the middle left 
		inferior edge of the bounding box (left). The ray intersects the mesh 
		at point $p_a$ that we consider the axilla center. We plot the 10 
		nearest vertices to $p_a$ (right). As it can be observed, the axilla is 
		properly recognized.}
	\label{fig:axilla-recognition}
\end{figure}

\subsection{Human Body Mesh Signature}\label{subsec:hbm_signature}
Intuitively, we would like to measure the chest circumference below the arms 
at the \textit{widest} part of the torso and the waist circumference at the 
\textit{narrowest} part beneath the chest but above the hips. This intuition is 
compliant 
with prior body dimensions standardized definition, for example, the European 
standard EN 13402-1 \cite{en13402-1}. Similarly, the pelvis 
circumference is measured often around the \textit{widest} part of hips and 
buttocks. The general idea is to cast the chest, waist and pelvis circumference 
estimation problem as an optimization problem. If we are able to establish a 
function that measures geodesics on the mesh, we can search for its maximum, 
e.g., chest in the chest region.

We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves\cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi_j}$ parallel (with normal $\vec{n}_{||}$) to the 
floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j \in 
\mathbb{R}$ along the y-axis. Since $\mathcal{M}$ is triangulated, the boundary 
of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the intersection boundary length as
\begin{equation}\label{eq:boundary_length}
\mathcal{BL}(\mathcal{M}, \vec{n}_{||}, q) = \sum_{i = 
	1}^{i = N}s_i
\end{equation}

@@

\begin{figure}[h!]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=0.5\linewidth]{subject_2_with_cutting_planes_xrays.png}
	\end{center}
	\caption{3D mesh iterative slicing. We 
		can use a plane $\boldsymbol{\pi_j}$ parallel (with normal 
		$\vec{n}_{||}$) to the 
		floor to intersect the mesh at point $q_j = (0, q^y_j, 0), \,  q^y_j 
		\in 
		\mathbb{R}$ along the y-axis. By varying point $q_j$ (we show here 40), 
		we can obtain 
		slices.}
	\label{fig:subjects_planes_xrays}
\end{figure}

Starting 
from the 
top $q_t$ of the bounding box we slice iteratively mesh $\mathcal{M}$ with 
plane 
$\boldsymbol{\pi_j}$ every $m$-meters along the y-axis until we reach the 
bounding box bottom 
$q_b$. Figure \ref{fig:subjects_planes_xrays} shows 40 of these 
slides. Next, we assemble the mesh sliding vector $\vec{\mathcal{L}}$. The 
elements of 
this vector are slice point y-coordinates

\begin{align}\label{eq:mesh_slice_vector}
\vec{\mathcal{L}}(\mathcal{M}, \vec{n}_{||}, m) = \left[ q_t, q_t+m, \cdots, 
q_b \right]
\end{align}

Finally, we apply equation \ref{eq:boundary_length} to every element of 
$\mathcal{L}$ in equation \ref{eq:mesh_slice_vector}. In other words, for every 
slice at point 
$q_j$, we compute $\mathcal{BL}(q_j)$. Here we drop from the notation mesh 
$\mathcal{M}$ and plane normal $\vec{n}_{||}$ because they remain constant. We 
then can 
define the \textbf{human body mesh signature} $\mathcal{MS}: 
\vec{\mathcal{L}} \to \mathbb{R}$ that maps every slice at point $q_j$ to the 
corresponding boundary length $\mathcal{BL}(q_j)$.

\begin{align}\label{eq:mesh_signature_short}
\mathcal{MS}(\vec{\mathcal{L}}) = [\mathcal{BL}(\mathcal{L}_1), 
\mathcal{BL}(\mathcal{L}_2), \cdots, 
\mathcal{BL}(\mathcal{L}_{|\vec{\mathcal{L}}|})]
\end{align}

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(\mathcal{M}, \vec{n}_{||}, m) = 
[\mathcal{BL}(q_t), \mathcal{BL}(q_t+m), \cdots, \mathcal{BL}(q_b)]
\end{align}

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Figure_1.eps}
	\end{center}
	\caption{Human Body Mesh Signature for male and female meshes. The 
		function resembles a rotated silhouette of the human body and exhibits 
		several \textit{extrema}. See section 
		\ref{subsec:mesh_signature_extrema} for discussion of these extrema.}
	\label{fig:hbm_signature}
\end{figure}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane parallel to the floor (with normal 
$\vec{n}_{||} = (0, 
1,0)$) using the library trimesh \cite{trimesh}. Note that the function as 
defined by equation \ref{eq:mesh_signature} is bounded and not continuous. It 
resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

In general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The chest circumference $cc$ is the local maximum within the chest 
	region.
	\item The waist circumference $wc$ (not to be confused with the natural 
	waist line) is the minimum within the waist region.	
	\item The pelvis circumference $pc$ is the maximum within the pelvis region.
\end{enumerate}

\begin{align}\label{eq:cc}
cc &= \mathcal{L}_i \in \mathcal{CR}\\
\mathcal{MS}(\mathcal{M}, \vec{n}_{||}, m) &= \mathcal{L}_i \in \mathcal{CR}
\end{align}
